Mapping Cancer: the Potential of Cartograms and Alternative Map Displays

### 2020-05-22 letter2/submission2.Rmd
Conclusions and recommendations

@@

**Reviewer suggestion:**

All minor comments including spell checking, grammar  were addressed in the process of substantial editing.

>>>>

**Reviewer suggestion:**
**Could you show an example? (of a completed contiguous cartogram that fully resolved the population transformation of areas)**
Sadly I could not find any packages in R that created a complete contiguous cartogram. Allowing the algorithm to run for further iterations saw the error rise, we chose the amount of iterations that led to the minimum error.

@@

**Reviewer suggestion:**
**Can you summarise these here? (suggestions and comments to help map creators best communicate their health data and spatial analysis)**

An additional paragraph has been included that dicusses the recommendations made by Bell et al. for choosing desirable alternative map diplays.

*Some mix of tiling, faceting or even micromaps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies. Bell et al. [@CPISACA] provide suggestions and comments to help map creators best communicate their health data and spatial analysis. The authors highlighted that the map design chosen should be tested on a representative sample of potential consumers, to ensure that the target audience is not misinformed by the display. The authors encourage the consideration of map types beyond the traditional classed choropleth map, but warn that sound cartographic principles must be employed to ensure effective communication to the public. A clear definition of the purpose of the display can help map creators to select the design that best communicates the statistic of interest [@CPISACA]. Table 4 lists several features, or limitations, of each alternative display in contrast to the commonly used choropleth map. The desirable features of each display can be contrasted within the table, this can be used to help inform the choice of map creators as they consider each alternative display. Map creators should choose a display that best communicates the statistic according to the purpose of their display.*

**Reviewer suggestion:**
**Not clear what “sometimes” means (in table 3).**
The use of Yes, No and Possible in the calls of Table 3 has been explained in the new caption:

*Summary of features and constraints of common mapping methods used to display cancer statistics. Yes means that the map dispay always employs the characteristic. No means that the display does not employ the characteristic. Possible means it is possible to employ the characteristic when using the display, but it is not necessary to produce the display.*

**Reviewer suggestion:**
**Suggest either a separate recommendations section or very clearly laying out the recommendations here. Do you recommend always (and only) using alternative visualisations, regardless of the audience, or are you suggesting supplying multiple forms of visualisations? Does your recommendation differ depending on the country/data? Ensure your final recommendations are clearly reflected in your abstract wording also. **

The following paragraph has been included to make clear the recommendations of the authors to the readers.

*Choropleth map displays are familiar and useful for map users to orient themselves to the statistic and additional features of the display. A choropleth map display may be the most effective display for the data considering budget, time, and maintenance contraints. Map creators should consider the shapes and discrepancies in size within the set of geographic units used for analysis, and whether the features of the dataset warrant transformation of the geographic units. Additional mapping methods should be considered by map creators during the development of a cancer atlas, as alternative displays may align better with the purpose of the display. Incorporating the population of the geographic units may better inform the map readers of the distribution of the statistic. The importance and potential impact of utilising a display that emphasises the underlying population should be considered during the design phase. Cartograms and tile maps should be adopted as primary displays going forward, or secondary displays if the familiarity of the choropleth map will benefit the users. Clear explanations of the interpretation of the geogephic units in alternative displays should also be included, along with the features of the distribution of the statistic. *

**Reviewer suggestion:**
**Isn’t the most ‘basic’ the number of cancers?**
We have removed the statement: The most basic is the incidence rate.

**Reviewer suggestion:**
**Crude or age standardized?**
We have clarified that the age standardized rates are of interest here.

**Reviewer suggestion:** XXX
I would suggest replacing this section with a discussion of relative versus absolute measures.

All minor comments including spell checking and grammar fixes were addressed in the process of editing.

### 2020-05-22 literature.Rmd
Conclusions and recommendations

Some mix of tiling, faceting or even micromaps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies. Moore and Carpenter [@SAMGIS] and Bell et al. [@CPISACA] provide suggestions and comments to help map creators best communicate their health data and spatial analysis.

Table 4: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics. Yes means that the map dispay always employs the characteristic. No means that the display does not employ the characteristic. Possible means it is possible to employ the characteristic when using the display, but it is not necessary or required.

@@

Monmonier [@HTLWM] says that interactivity can be used to allow users to explore the map for more information and provides flexibility for the display. The user can toggle between different variables, map views or even multiple realizations of future scenarios [@goodchild1994introduction]. This provides additional mechanisms for the users to digest the uncertainty of the available information [@maceachren1992visualizing, @van1994visualization]. When the needs of the audience are changeable and are also the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions. This can allow inspection of the data from many views [@DQBCM]. User interaction with maps helps to understand and interpret the spatial distribution of disease, to validate, explain or explore the presented statistics and their relationships to each other [@TNTEA].

@@

The Australian Cancer Atlas [@TACA] provides [tours](
https://atlas.cancer.org.au/app/tour/lungcancer) that change the display to draw users' attention to areas on the map that are relevant to the story.
This implementation of animation gives users tools to plan their exploration.

@@

<!-- summary of mapping -->
This paper provides an overview of mapping practices as commonly used for cancer atlases and recommends new approaches, such as cartograms and hexagon tile maps that should be adopted going forward. The conventional approach is the choropleth map, and it is widely used. The choropleth map suffers when there are small geographic units, as occurs in Australia where the population is concentrated on the coast, the information about the burden of cancer on those communities can be hidden. Making an inset can clarify congested regions but this breaks the viewers' attention as they shift focus from the map to the inset, and if there are many congested areas, many insets would be needed. The map alternatives implement trade-offs between the familiar shapes, and the importance of the geographic areas in the context of the areas. Given the population or a cancer statistic for each area, the geographic size or shape will change. Alternative displays allow the spatial distribution of cancer data to be digested by map users.

<!-- statistics -->
Many statistics are commonly used in cancer displays. The most basic is the incidence rate. It is common to see relative rates which measure how far a region is above or below the average. The purpose of using a relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates. 
<!--However, we lose the incidence rate information and thus interpretability. --> A region might be much higher than average, but it may not be close to a health concern, because all regions have a low incidence. Supplementary materials can allow map users to recognise when this occurs.

<!-- interaction -->
Interaction with maps is an important component of public atlases, and is easy to add with today's technology. The purpose is to provide access to more information than is possible to display in a single map, without overwhelming the viewer. Too many choices can similarly overwhelm a viewer, and thus decisions do need to be made about content to provide for accurate and comprehensive communication of information. Similarly, providing ways for users to interact with the display encourages engagement, and creative, efficient, elegant, interactive tools elicit curiosity about the data.

>>>>

Some mix of tiling, faceting or even micromaps, which allow some spatial continuity while also zooming into small areas, are good solutions for difficult geographies. Bell et al. [@CPISACA] provide suggestions and comments to help map creators best communicate their health data and spatial analysis. The authors highlighted that the map design chosen should be tested on a representative sample of potential consumers, to ensure that the target audience is not misinformed by the display. The authors encourage the consideration of map types beyond the traditional classed choropleth map, but warn that sound cartographic principles must be employed to ensure effective communication to the public.
A clear definition of the purpose of the display can help map creators to select the design that best communicates the statistic of interest [@CPISACA]. Table 3 lists several features, or limitations, of each alternative display in contrast to the commonly used choropleth map. The desirable features of each display can be contrasted within the table, this can be used to help inform the choice of map creators as they consider each alternative display. Map creators should choose a display that best communicates the statistic according to the purpose of their display.

Table 3: <!--\label{tab:methods}--> Summary of features and constraints of common mapping methods used to display cancer statistics. Yes means that the map dispay always employs the characteristic. No means that the display does not employ the characteristic. Possible means it is possible to employ the characteristic when using the display, but it is not necessary to produce the display.

@@

Monmonier [@HTLWM] recommends using interactivityto allow users to explore the map for more information and provide flexibility for the display. The user can toggle between different variables, map views or even multiple realizations of future scenarios [@goodchild1994introduction]. This provides additional mechanisms for the users to digest the uncertainty of the available information [@maceachren1992visualizing, @van1994visualization]. When the needs of the audience are changeable and are also the priority, the map creator can allow interactivity for map users to explore a data set through dynamic interactions. This can allow inspection of the data from many views [@DQBCM]. User interaction with maps helps to understand and interpret the spatial distribution of disease, to validate, explain or explore the presented statistics and their relationships to each other [@TNTEA].

@@

The Australian Cancer Atlas [@TACA] provides [tours](
https://atlas.cancer.org.au/app/tour/lungcancer) that change the display to draw users' attention to areas on the map that are relevant to the interpretation of the statistic displayed.
This implementation of animation gives users tools to plan their exploration.

@@

<!-- summary of mapping -->
This paper provides an overview of mapping practices as commonly used for cancer atlases. The conventional approach is the choropleth map, and it is widely used. The choropleth map suffers when there are small geographic units, as occurs in Australia where the population is concentrated on the coast, the information about the burden of cancer on those communities can be hidden. Making an inset can clarify congested regions but this breaks the viewers' attention as they shift focus from the map to the inset, and if there are many congested areas, many insets would be needed. The map alternatives implement trade-offs between the familiar shapes, and the importance of the geographic areas in the context of the areas. Given the population or a cancer statistic for each area, the geographic size or shape will change. Alternative displays allow the spatial distribution of cancer data to be digested by map users.

Choropleth map displays are familiar and useful for map users to orient themselves to the statistic and additional features of the display. A choropleth map display may be the most effective display for the data considering budget, time, and maintenance contraints. Map creators should consider the shapes and discrepancies in size within the set of geographic units used for analysis, and whether the features of the dataset warrant transformation of the geographic units.
Additional mapping methods should be considered by map creators during the development of a cancer atlas, as alternative displays may align better with the purpose of the display. Incorporating the population of the geographic units may better inform the map readers of the distribution of the statistic. The importance and potential impact of utilising a display that emphasises the underlying population should be considered during the design phase.
Cartograms and tile maps should be adopted as primary displays going forward, or secondary displays if the familiarity of the choropleth map will benefit the users. Clear explanations of the interpretation of the geogephic units in alternative displays should also be included, along with the features of the distribution of the statistic.

<!-- statistics -->
Many statistics are commonly used in cancer displays. It is common to see incidence rates, or relative rates which measure how far a region is above or below the average. The purpose of using an age standardised relative rate is, perhaps the desire to pinpoint the areas that need attention because they have higher than expected rates. 
<!--However, we lose the incidence rate information and thus interpretability. --> A region might be much higher than average, but it may not be close to a health concern, because all regions have a low incidence. Supplementary materials can allow map users to recognise when this occurs.

<!-- interaction -->
Interaction with maps is an important component of public atlases, and is becoming increasingly straightforward to add with today's technology. The purpose of interaction in public atlasses is to provide access to more information than is possible to display in a single map, without overwhelming the viewer. Too many choices can similarly overwhelm a viewer, and thus decisions do need to be made about content to provide for accurate and comprehensive communication of information. Similarly, providing ways for users to interact with the display encourages engagement, and creative, efficient, elegant, interactive tools elicit curiosity about the data.
